---
title: "Project1"
author: "Xingli Ma"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  github_document:
    toc: yes
    toc_depth: 3
---       

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, error = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
```         

The following R packages are reqired to run the code to create this vignette.    

```{r}
require(rmarkdown)
require(knitr)
require(dplyr)
require(tidyverse)
require(ggplot2)
require(DT)
library(jsonlite)
library(httr)
```             

## Introduction

I created a vignette for reading and summarizing data from the National Hockey League’s
(NHL) API. Three main components are present in this vignette: extracting tables and querying information from NFL Records API, extracting tables and querying information from NFL Stats API, and exploratory data analysis.

## NFL Records API and NFL Stats API 

### Retriving Tables and Quering Information from NFL Records API   

Create a wrapper function to return the records for a given team ID or franchise ID.    

```{r}
team_Record <- function(teamId_OR_franchiseId, ...){
  
  # construct URLs
  full_url_records <- function(table_name, ...){
  base_url <- "https://records.nhl.com/site/api"
  return(paste0(base_url, table_name, ...))
}
  franchise <- full_url_records("/franchise?cayenneExp=mostRecentTeamId=", teamId_OR_franchiseId)
  franchise_teams <- full_url_records("/franchise-team-totals?cayenneExp=teamId=", teamId_OR_franchiseId)
  franchise_season <- full_url_records("/franchise-season-records?cayenneExp=franchiseId=", teamId_OR_franchiseId)
  goalie <- full_url_records("/franchise-goalie-records?cayenneExp=franchiseId=", teamId_OR_franchiseId)
  skater <- full_url_records("/franchise-skater-records?cayenneExp=franchiseId=", teamId_OR_franchiseId)
  franchise_detail <- full_url_records("/franchise-detail?cayenneExp=mostRecentTeamId=", teamId_OR_franchiseId)
  
 # retrieve the NHL records information and convert it to data frames  
  franchise_record <- jsonlite::fromJSON(content(GET(franchise),"text"),       flatten=TRUE)$data
  franchise_teams_record <- jsonlite::fromJSON(content(GET(franchise_teams),"text"), flatten=TRUE)$data
  franchise_season_record <- jsonlite::fromJSON(content(GET(franchise_season),"text"), flatten=TRUE)$data
  goalie_record <- jsonlite::fromJSON(content(GET(goalie),"text"), flatten=TRUE)$data
  skater_record <- jsonlite::fromJSON(content(GET(skater),"text"), flatten=TRUE)$data
  franchise_detail_record <- jsonlite::fromJSON(content(GET(franchise_detail),"text"), flatten=TRUE)$data 
  return(list(df1_franchise=franchise_record, df2_teams=franchise_teams_record,   df3_season=franchise_season_record, df4_goalie=goalie_record, df5_skater=skater_record, df6_adminDetail=franchise_detail_record))
}

# Query NHL records information according to the team ID or franchise ID given by a user
team_Record(2)
```  


### Retriving Tables and Quering Information from NFL Stats API   

Create a wrapper function to return the stats for a given team ID.   

```{r}
team_Stats <- function(teamID, ...) {
  # construct URLs
  full_url_stats <- function(teamID){
  base_url <- "https://statsapi.web.nhl.com/api/v1/teams/"
  tab_name <- "/?expand=team.stats"
  return(paste0(base_url, teamID, tab_name))
  }

  # retrieve the NHL stats information and convert it to data frames
  NHLstats <- function(full_url_stats){
  return(jsonlite::fromJSON(content(GET(full_url_stats),"text"), flatten=TRUE))
  }
  
  teamStatsDf <- NHLstats(full_url_stats(teamID))$teams[[8]][[1]]
  teamStatsDf$splits[[1]]
}

# Query NHL stats information according to the team ID or franchise ID given by a user
team_Stats(1)
```   


## Exploratory Data Analysis   

### Tables Retrieved from NHL Records API   

```{r}
library(jsonlite)
library(httr)
# construct URLs
full_url_records <- function(tabl_name){
  base_url <- "https://records.nhl.com/site/api"
  return(paste0(base_url, tabl_name))
}
franchise <- full_url_records("/franchise")
franchise_totals <- full_url_records("/franchise-team-totals")
franchise_season <- full_url_records("/franchise-season-records")
goalie <- full_url_records("/franchise-goalie-records")
skater <- full_url_records("/franchise-skater-records")
admin <- full_url_records("/franchise-detail")

# retrieve the NHL records information and convert it to data frames
NHLrecords <- function(full_url_records){
  return(jsonlite::fromJSON(content(GET(full_url_records),"text"), flatten=TRUE))
}
franchiseDf <- NHLrecords(franchise)$data
franchise_totalsDf <- NHLrecords(franchise_totals)$data
franchise_seasonsDf <- NHLrecords(franchise_season)$data
goalieDf <- NHLrecords(goalie)$data
skaterDf <- NHLrecords(skater)$data
adminDf <- NHLrecords(admin)$data
```   

```{r}
#franchiseDf
#franchise_totalsDf
#franchise_seasonsDf
#goalieDf
#skaterDf
#adminDf
```   

### Create Plots for NHL Records  

Bar plot on goalie and skater information.  

```{r}
dim(goalieDf)
dim(skaterDf)
goalieDf_new <- cbind(goalieDf,df=rep(1,1078))
skaterDf_new <- cbind(skaterDf,df=rep(2,17209))

goalie_skater <- bind_rows(goalieDf_new, skaterDf_new)
dim(goalie_skater)
goalie_skater_new <- filter(goalie_skater, gameTypeId == 2, activePlayer == TRUE)
goalie_skater_new
```   

```{r franchise_totals bar plot}
g <- ggplot(data=franchise_totalsDf, aes(x=gameTypeId))
g + geom_bar(aes(fill=wins)) +
  labs(x="game Type Id")
```       


Histogram plot on goalie and skater information.     

```{r}

```    


Box plot on goalie and skater information.     

```{r}

```    


Scatter plot on goalie and skater information.     

```{r}

```    














